---
title: "Fancy axis plots!"
output:
  html_document: default
  html_notebook: default
---

## First, we need to get the proper libraries, many of which are dev versions from github:
```{r,warning=F, message=F}
#install.packages('devtools')
#library(devtools)
#install_github("clauswilke/ggjoy")
#install_github("wilkelab/cowplot")
#install.packages('ggstance')

library(ggplot2)
library(ggstance)
library(ggjoy)
library(cowplot)
```

With that done, lets try some examples:

```{r}
pmain = ggplot(iris,aes(x=Sepal.Length,y=Sepal.Width, color=Species)) + geom_point()

xdens = axis_canvas(pmain,axis='x') +
  geom_ridgeline(data=iris,aes(x=Sepal.Length,y=0,height=..density..,fill=Species),
                 stat='xdensity',alpha=0.7,size=0.2,trim=F)

ydens = axis_canvas(pmain,axis='y') +
  geom_vridgeline(data=iris,aes(y=Sepal.Width,x=0,width=..density..,fill=Species),
                 stat='ydensity',alpha=0.7,size=0.2,trim=F)

p1 = insert_xaxis_grob(pmain,xdens, grid::unit(0.2, "null"), position='top')
p2 = insert_yaxis_grob(p1, ydens, grid::unit(.2, "null"), position = 'right')

ggdraw(p2)
```

Neat! Lets unpack this a bit more:

We are starting with a basic ggplot-based scatter plot on the `iris` dataset. This is defined as `pmain`

```{r}
pmain = ggplot(iris,aes(x=Sepal.Length,y=Sepal.Width, color=Species)) + geom_point()
```
From here, we are using the `axis_canvas` function from `cowplot`. This takes an existing ggplot2 plot and creates a new plot for the axis. It allows for custom 'axis-like' objects or annotations in the margins. Check the help for details on this.

We also are using the `geom_ridgeline` and `geom_vridgeline` functions from `ggjoy`. These per the description: 

'Plots the sum of the y and height aesthetics versus x, filling the area between y and y + height with a color'

So we are gathering the datapoints present at every x value and creating a density plot. Reverse the concept for vridgeline.

We then create our x and y density plots:
```{r}
xdens = axis_canvas(pmain,axis='x') + #define the canvas from pmain on the x axis
  geom_ridgeline(data=iris, #geom_ridgeline is from the ggjoy package
                 aes(x=Sepal.Length, #dataset
                     y=0, #not used
                     height=..density.., #the height of the ridgeline, we are not providing 'identity' but rather density
                     fill=Species), #color fill by Species factor
                 stat='xdensity', #the statistic used
                 alpha=0.7, #a bit transparent
                 size=0.2, #fairly small binding line for the density curves
                 trim=F) #do not truncate the density curve to the range of the data

ydens = axis_canvas(pmain,axis='y') +
  geom_vridgeline(data=iris,aes(y=Sepal.Width,x=0,width=..density..,fill=Species),
                 stat='ydensity',alpha=0.7,size=0.2,trim=F)
```

These axis plots are truly plots of their own and can be viewed individually:
```{r}
ggdraw(xdens)
```

From here, we have all the pieces and can build our final version of the plot. We now are using the other functions of cowplot - `insert_xasis_grob` and `insert_yaxis_grob` - to pull it all together:

```{r}
p1 = insert_xaxis_grob(plot=pmain, #the main plot to insert into 
                       grob=xdens, #the 'grob' (grid graphical object) you want to insert for the axis
                       height=grid::unit(0.2, #vector size for how high you want it to be
                                         units="null"), #we are not working in cm, in, or anything like that
                       position='top')

#now tack on the yaxis plot to p1:
p2 = insert_yaxis_grob(p1, ydens, grid::unit(.2, "null"), position = 'right')

#plot it!
ggdraw(p2)
```
